---
title:  "P8105 Data Science: Homework 1"
author: "Annie Clark (alc2279)"
date:   2018-09-15
output: html_document
---

#Problem 0.1 - Workflow

###_Prompt:_
This “problem” focuses on the use of R Markdown to write reproducible reports,
GitHub for version control, and R Projects to organize your work.

To that end:

* create a public GitHub repo + local R Project; we suggest naming this 
repo / directory p8105_hw1_YOURUNI (e.g.  p8105_hw1_ajg2202 for Jeff), but that’s
not required
* create a single .Rmd file named p8105_hw1_YOURUNI.Rmd that renders to 
github_document

Your solutions to Problems 1 and 2 should be implemented in your .Rmd file, and 
your git commit history should reflect the process you used to solve these Problems.

For this Problem, we will assess adherence to the instructions above regarding repo
structure, git commit history, and whether we are able to knit your .Rmd to ensure
that your work is reproducible._

#Problem 0.2 - Styling

This “problem” focuses on correct styling for your solutions to Problems 1 and 2.
We will look for:

* meaningful variable / object names
* readable code (one command per line; adequate whitespace and indentation; etc)
* clearly-written text to explain code and results
* a lack of superfluous code (e.g. no unused variables are defined)

#Problem 1 - Variable types and coercion

##Part 1 - Create  a data frame
Create a data frame comprised of:

* A random sample of size 10 from a uniform [0, 5] distribution
* A logical vector indicating whether elements of the sample are greater than 2
* A (length-10) character vector
* A (length-10) factor vector

```{r}

library(tidyverse)

## set seed to ensure reproducibility
set.seed(2)

##define datframe containing four different types of vectors
prob_1_data_frame = tibble(
  vec_num_random_sample = rnorm(10, sd = 5),
  vec_logical_random_sample_greater_2 = vec_num_random_sample > 2,
  vec_char = c("Puppies", "make", "the", "world", "go", "round", "and", "round", 
               "and", "round"),
  vec_factor = factor(c("male", "male", "female", "female", "male", "male", 
                        "female", "female", "male", "female"))
)
```


##Part 2 - Mean of variable types
Try to take the mean of each variable in your dataframe. What works and what 
doesn’t? Why?


```{r}

## specify to use this data frame when evaluating variables
attach(prob_1_data_frame)

## mean of vec_num_random_sample
mean(vec_num_random_sample)

##mean of vec_logical_random_sample_greater_2
mean(vec_logical_random_sample_greater_2)

## mean of vec_char
mean(vec_char)

## mean of vec_factor
mean(vec_factor)

```

The numeric 

##Part 3 - Coercing variables

In some cases, you can explicitly convert variables from one type to another. 
Write a code chunk that applies the as.numeric function to the logical, character,
and factor variables (please show this chunk but not the output). What happens? 


```{r}

vec_random_greater_2_factor = as.factor(prob_1_data_frame$vec_num_random_sample) 
class(vec_random_greater_2_factor)

vec_char_factor = as.factor(vec_char)
class(vec_char_factor)

vec_factor_factor = as.factor(vec_factor) 
class(vec_factor_factor)

```

##Part 4 - Coercing variables (cont.)

In a second code chunk, convert your character variable from character to factor
to numeric; similarly, convert your factor variable from factor to character to 
numeric. What happens?

```{r}
#vec_char: character to factor to numeric
class(vec_char)
vec_char
#character

vec_char_factor2 = as.factor(vec_char)
class(vec_char_factor2)
vec_char_factor2
#factor

vec_char_num = as.numeric(vec_char_factor2)
class(vec_char_num)
#numeric
vec_char_num

## orders them; lose the character information

#vec_factor: factor to character to numeric

class(vec_factor)
#factor
vec_factor

vec_factor_char = as.character(vec_factor)
class(vec_factor_char)
#character 
vec_factor_char

vec_factor_char_num = as.numeric(vec_factor_char)
#NAs introduced by coercion 
class(vec_factor_char_num)
vec_factor_char_num
#loses all information => goes to NAs

```



#Problem 2 - Plotting and inline R code

##Part 1 - Create a Data Frame
Create a data frame comprised of:

* x: a random sample of size 1000 from a standard Normal distribution
* y: a random sample of size 1000 from a standard Normal distribution
* A logical vector indicating whether the x + y > 0
* A numeric vector created by coercing the above logical vector
* A factor vector created by coercing the above logical vector

```{r}

## set seed to ensure reproducibility
set.seed(9) 

##define datframe containing 2 random sample variables. Next, a logical variable
##is created based on the addition of the random sample variables, and is then 
##coerced into a numeric and factor variable. 
prob_2_data_frame = tibble(
  x = rnorm(1000),
  y = rnorm(1000),
  z_logical = (x + y) > 0,
  z_num = as.numeric(z_logical),
  z_factor = as.factor(z_logical)
)

```

##Part 2 - Inline Description of Data Frame Variables

* Write a short description of your vector using inline R code, including: 
* the size of the dataset * the mean and median of x * the proportion of cases 
for which the logical vector is TRUE

Next, I am going to describe the dataset that I just created.  There 
are 'r nrow(example_df2)' columns and 'r nrow(example_df2)' rows. The mean 
of x is 'r mean(x)' and the median is 'r median(x)'. There are XX cases in 
which the logical vector, z is TRUE. 
##still need to figure out last inline code

##Part 3 - Scatterplots

Make a scatterplot of y vs x; color points using the logical variable (adding 
color = ... inside of aes in your ggplot code should help). Make a second and 
third scatterplot that color points using the numeric and factor variables, 
respectively, and comment on the color scales.

```{r}

##scatterplot of y vs x using z_logical to color points
ggplot(prob_2_data_frame, aes(x = y, y = x, color = z_logical)) + geom_point()

##scatterplot of y vs x using z_num to color points
ggplot(prob_2_data_frame, aes(x = y, y = x, color = z_num)) + geom_point()

##scatterplot of y vs x using z_factor to color points
ggplot(prob_2_data_frame, aes(x = y, y = x, color = z_factor)) + geom_point()

```

##Part 4 - Exporting Scatterplot

Export your first scatterplot to your project directory using ggsave.

```{r}
##scatterplot of y vs x using z_logical to color points
ggplot(prob_2_data_frame, aes(x = y, y = x, color = z_logical)) + geom_point()

##scatterplot of y vs. x using z_logical to color points is exported to project
##diretory
ggsave("prob_2_random_sample_scatter_plot.pdf", height = 4, width = 6)

```

