---
title:  "P8105 Data Science: Homework 1"
author: "Annie Clark (alc2279)"
date:   2018-09-15
output: github_document
---

#Problem 0.1

This “problem” focuses on the use of R Markdown to write reproducible reports,
GitHub for version control, and R Projects to organize your work.

To that end:

* create a public GitHub repo + local R Project; we suggest naming this 
repo / directory p8105_hw1_YOURUNI (e.g.  p8105_hw1_ajg2202 for Jeff), but that’s
not required
* create a single .Rmd file named p8105_hw1_YOURUNI.Rmd that renders to 
github_document

Your solutions to Problems 1 and 2 should be implemented in your .Rmd file, and 
your git commit history should reflect the process you used to solve these Problems.

For this Problem, we will assess adherence to the instructions above regarding repo
structure, git commit history, and whether we are able to knit your .Rmd to ensure 
that your work is reproducible.

#Problem 0.2

This “problem” focuses on correct styling for your solutions to Problems 1 and 2.
We will look for:

* meaningful variable / object names
* readable code (one command per line; adequate whitespace and indentation; etc)
* clearly-written text to explain code and results
* a lack of superfluous code (e.g. no unused variables are defined)

#Problem 1

Create a data frame comprised of:

* A random sample of size 10 from a uniform [0, 5] distribution
* A logical vector indicating whether elements of the sample are greater than 2
* A (length-10) character vector
* A (length-10) factor vector

```{r}

library(tidyverse)

set.seed(1)

example_df = tibble(
  vec_random = rnorm(10, mean = 2.5, sd = 1), ##still need to figure out what to set SD
  vec_random_greater_2 = vec_random > 2,
  vec_char = c("Puppies", "make", "the", "world", "go", "round", "and", "round", 
               "and", "round"),
  vec_factor = factor(c("male", "male", "female", "female", "male", "male", 
                        "female", "female", "male", "female"))
)

```

Try to take the mean of each variable in your dataframe. What works and what 
doesn’t? Why?



```{r}
mean(vec_random)

mean(vec_random_greater_2)

mean(vec_char)

mean(vec_factor)

```

In some cases, you can explicitly convert variables from one type to another. 
Write a code chunk that applies the as.numeric function to the logical, character, 
and factor variables (please show this chunk but not the output). What happens? 


```{r}

vec_random_greater_2_factor = as.factor(vec_random_greater_2) 
class(vec_random_greater_2_factor)

vec_char_factor = as.factor(vec_char)
class(vec_char_factor)

vec_factor_factor = as.factor(vec_factor) 
class(vec_factor_factor)

```


In a second code chunk, convert your character variable from character to factor
to numeric; similarly, convert your factor variable from factor to character to 
numeric. What happens?

```{r}
#vec_char: character to factor to numeric
class(vec_char)
vec_char
#character

vec_char_factor2 = as.factor(vec_char)
class(vec_char_factor2)
vec_char_factor2
#factor

vec_char_num = as.numeric(vec_char_factor2)
class(vec_char_num)
#numeric
vec_char_num

## orders them; lose the character information

#vec_factor: factor to character to numeric




```



#Problem 2

Create a data frame comprised of:

* x: a random sample of size 1000 from a standard Normal distribution
* y: a random sample of size 1000 from a standard Normal distribution
* A logical vector indicating whether the x + y > 0
* A numeric vector created by coercing the above logical vector
* A factor vector created by coercing the above logical vector
* Write a short description of your vector using inline R code, including: 
* the size of the dataset * the mean and median of x * the proportion of cases 
for which the logical vector is TRUE

```{r}

```

Make a scatterplot of y vs x; color points using the logical variable (adding 
color = ... inside of aes in your ggplot code should help). Make a second and 
third scatterplot that color points using the numeric and factor variables, 
respectively, and comment on the color scales.

```{r}

```

Export your first scatterplot to your project directory using ggsave.

```{r setup, include=FALSE}

```

